Pillar 1 - Sequential Pattern Mining

### 2019-02-05 Pillar 1 - Sequential Pattern Mining.Rmd
rename item

```{r AR data prep 1}

@@

### Adding not taken courses
```{r transactions not taken}
d_transactions_not_taken <- expand.grid(

@@

    taken = case_when(
      is.na(Grade) ~ "not taken",
      TRUE         ~ "taken"
      ),

    taken_grade = case_when(
      is.na(Grade) ~ "not taken",
      Grade < 5.5  ~ "fail",
      TRUE         ~ "pass"
      ),

    item_taken       = paste(item, taken      , sep = "_"),
    item_taken_grade = paste(item, taken_grade, sep = "_")

@@

```
###Expanded Transcripts
```{r transactions expanded}

@@

  # Arrange per sequenceID and item for making sequences
  arrange(
    sequenceID,
    item
    ) %>%

  # Identifying item
  mutate(
    item_Grade = paste(item, Grade_round, sep = "_")
    )
```

@@

# Probability of failing, having a low grade of a grade less than x in a course
d_support <- d_transactions %>%
  mutate(
    grade_round = round(Grade)
  ) %>%
  group_by(
    item
    ) %>%
  summarise(
    rate.fail = mean(PF == "fail"),
    rate.low  = mean(HL == "low"),
    rate.0    = mean(grade_round <= 0),
    rate.1    = mean(grade_round <= 1),
    rate.2    = mean(grade_round <= 2),
    rate.3    = mean(grade_round <= 3),
    rate.4    = mean(grade_round <= 4),
    rate.5    = mean(grade_round <= 5),
    rate.6    = mean(grade_round <= 6),
    rate.7    = mean(grade_round <= 7),
    rate.8    = mean(grade_round <= 8),
    rate.9    = mean(grade_round <= 9),
    rate.10   = mean(grade_round <= 10)
    )

@@

### Transactions for Apriori
```{r making transactions}

@@

#
# Making transactions
transactions                  <- make_transaction(item = item   )
transactions_PF               <- make_transaction(item = item_PF)
transactions_HL               <- make_transaction(item = item_HL)
transactions_Taken            <- make_transaction(data = d_transactions_not_taken, item = item_taken)
transactions_expanded_Grade   <- make_transaction(data = d_transactions_expanded, item = item_Grade)

@@

  sequences@itemsetInfo <- select(
    data,
    sequenceID,
    eventID
    ) %>%
    as.data.frame()

@@

#
# Making sequences
sequences          <- make_sequence(item = item   )
sequences_PF       <- make_sequence(item = item_PF)
sequences_HL       <- make_sequence(item = item_HL)

sequences_expanded_Grade  <- make_sequence(data = d_transactions_expanded, item = item_Grade)

#
# Checking sequences
# inspect(head(sequences, 10))

@@

  data %>%
    apriori(
      parameter = list(

        # Chosen parameters
        supp = 0,    # min support
        smax = 1,    # max support
        conf = 0,    # min confidence
        minlen = 2,  # min length of 
        maxlen = 2,  # max length of rule
        ext = TRUE,
        arem = "diff",
        aval = TRUE,
        minval = 0
        ),

@@

      control = list(
        verbose = FALSE
      )

@@

    # Exclude rules with support of `0`
    filter(
      count >= 1
    ) %>%

    # Clean variable `rules`
    mutate(
      rules = str_remove_all(rules, pattern = "[{]"),
      rules = str_remove_all(rules, pattern = "[}]")
    ) %>%
    separate(rules, into = c("lhs", "rhs"), sep = " => ") %>%

    # Editing output
    mutate_at(
      c("support", "lhs.support", "confidence", "lift"),
      funs(round(., 5))
    ) %>%
    select(
      lhs, rhs,
      support, count, lhs.support,
      confidence,
      lift
    ) %>%
    arrange(
      desc(count)
    )

@@

    # include lhs.support
    mutate(
      lhs_course = str_remove_all(lhs, pattern = "_fail|_low|_not taken")
      ) %>%
    left_join(
      select(data_support, item, !!enquo(type_rule)),
      by = c("lhs_course" = "item")
      ) %>%
    mutate(
      lhs.support = round(!!type_rule_enquo, 5)
      )  %>%
    select(
      - !!type_rule_enquo
      ) %>%

    # include rhs.support
    mutate(
      rhs_course = str_remove_all(rhs, pattern = "_fail|_low|_not taken")
      ) %>%
    left_join(
      select(data_support, item, !!type_rule_enquo),
      by = c("rhs_course" = "item")
      ) %>%
    mutate(
      rhs.support = round(!!type_rule_enquo, 5)
      ) %>%
    select(
      - !!type_rule_enquo
      ) %>%

    # order variable
    select(
      lhs_course, lhs, rhs, support, count, confidence, rhs.support, lift, lhs.support
    )

@@

#
# AR fail
AR_PF <- transactions_PF %>%
  make_AR(
    appearance = list(

      # impose lhs and rhs to be fail
      both = d_transactions %>%
        filter(PF == "fail") %>%
        distinct(`item_PF`) %>%
        pull
    ),

@@

#
# AR low
AR_PF <- transactions_HL %>%
  make_AR(
    appearance = list(

      # impose lhs and rhs to be low
      both = d_transactions %>%
        filter(HL == "low") %>%
        distinct(`item_HL`) %>%
        pull
    ),

@@

#
# AR not taken
AR_not_taken <- transactions_Taken %>%
  make_AR(
    appearance = list(

      # impose lhs and rhs to be not taken
      both = d_transactions_not_taken %>%
        filter(taken == "not taken") %>%
        distinct(`item_taken`) %>%
        pull
    ),

>>>>

### take, fail, low grade
```{r AR data prep 1}

@@

### take / not take, pass / fail / not take
```{r transactions not taken}
d_transactions_not_taken <- expand.grid(

@@

    TNT = case_when(
      is.na(Grade) ~ "not taken",
      TRUE         ~ "taken"
      ),

    TPF = case_when(
      is.na(Grade) ~ "not taken",
      Grade < 5.5  ~ "fail",
      TRUE         ~ "pass"
      ),

    item_TNT = paste(item, TNT, sep = "_"),
    item_TPF = paste(item, TPF, sep = "_")

@@

### Grade
```{r transactions expanded}

@@

  # Identifying item
  mutate(
    item_G = paste(item, Grade_round, sep = "_")
    )
```

@@

# Probability of failing, having a low grade, having a grade less than x per course
d_support <- d_transactions %>%
  mutate(
    grade_round = round(Grade)
  ) %>%
  group_by(
    item
    ) %>%
  summarise(
    rate.fail = mean(PF == "fail"),
    rate.low  = mean(HL == "low"),
    rate.0    = mean(grade_round <= 0),
    rate.1    = mean(grade_round <= 1),
    rate.2    = mean(grade_round <= 2),
    rate.3    = mean(grade_round <= 3),
    rate.4    = mean(grade_round <= 4),
    rate.5    = mean(grade_round <= 5),
    rate.6    = mean(grade_round <= 6),
    rate.7    = mean(grade_round <= 7),
    rate.8    = mean(grade_round <= 8),
    rate.9    = mean(grade_round <= 9),
    rate.10   = mean(grade_round <= 10)
    )

@@

### Transactions for Apriori
```{r making transactions, message = FALSE}

@@

#
# Making transactions
transactions     <- make_transaction(item = item   )
transactions_PF  <- make_transaction(item = item_PF)
transactions_HL  <- make_transaction(item = item_HL)
#transactions_TNT <- make_transaction(data = d_transactions_not_taken, item = item_TNT)
transactions_TPF <- make_transaction(data = d_transactions_not_taken, item = item_TPF)
transactions_G   <- make_transaction(data = d_transactions_grade    , item = item_G  )

@@

  sequences@itemsetInfo <- select(
    data,
    sequenceID,
    eventID
    ) %>%
    as.data.frame

@@

#
# Making sequences
sequences     <- make_sequence(item = item   )
sequences_PF  <- make_sequence(item = item_PF)
sequences_HL  <- make_sequence(item = item_HL)
#sequences_TNT <- make_sequence(data = d_transactions_not_taken, item = item_TNT) # compute ad hoc
#sequences_TPF <- make_sequence(data = d_transactions_not_taken, item = item_TPF) # compute ad hoc
sequences_G   <- make_sequence(data = d_transactions_grade    , item = item_G  )

#
# Checking sequences
# inspect(head(sequences_G, 10))

@@

  data %>%
    apriori(

      # Chosen parameters
      parameter = list(
        supp = 0,    # min support
        smax = 1,    # max support
        conf = 0,    # min confidence
        minlen = 2,  # min length of 
        maxlen = 2,  # max length of rule
        ext = TRUE,
        arem = "diff",
        aval = TRUE,
        minval = 0
        ),

@@

      # no printing during execution
      control = list(
        verbose = FALSE
        )

@@

    # Exclude rules with no support
    filter(
      count >= 1
      ) %>%

    # Clean variable `rules`
    mutate(
      rules = str_remove_all(rules, pattern = "[{]"),
      rules = str_remove_all(rules, pattern = "[}]")
      ) %>%
    separate(
      rules,
      into = c("lhs", "rhs"), 
      sep  = " => "
      )

@@

    # include lhs.support
    mutate(
      lhs_course = str_remove_all(lhs, pattern = "_fail|_low|_not taken")
      ) %>%
    left_join(
      select(data_support, item, !!type_rule_enquo),
      by = c("lhs_course" = "item")
      ) %>%
    mutate(
      lhs.support = !!type_rule_enquo
      )  %>%
    select(
      - !!type_rule_enquo
      ) %>%

    # include rhs.support
    mutate(
      rhs_course = str_remove_all(rhs, pattern = "_fail|_low|_not taken")
      ) %>%
    left_join(
      select(data_support, item, !!type_rule_enquo),
      by = c("rhs_course" = "item")
      ) %>%
    mutate(
      rhs.support = !!type_rule_enquo
      ) %>%
    select(
      - !!type_rule_enquo
      ) %>%

    # Editing output
    mutate_at(
      c("support", "lhs.support", "confidence", "rhs.support", "lift"),
      funs(round(., 5))
    ) %>%
    select(
      lhs, rhs,
      support, count, lhs.support,
      confidence, rhs.support,
      lift
    ) %>%
    arrange(
      desc(count)
    )

@@

#
# AR fail
AR_PF <- transactions_PF %>%
  make_AR(

    # impose lhs and rhs to be fail
    appearance = list(
      both = d_transactions %>%
        filter(PF == "fail") %>%
        distinct(item_PF) %>%
        pull
    ),

@@

#
# AR low
AR_PF <- transactions_HL %>%
  make_AR(

    # impose lhs and rhs to be low
    appearance = list(
      both = d_transactions %>%
        filter(HL == "low") %>%
        distinct(`item_HL`) %>%
        pull
    ),

@@

#
# AR not taken
AR_not_taken <- transactions_Taken %>%
  make_AR(

    # impose lhs and rhs to be not taken
    appearance = list(
      both = d_transactions_not_taken %>%
        filter(TNT == "not taken") %>%
        distinct(`item_TNT`) %>%
        pull
    ),
